---
title: "AKI"
author: "Thomas Oates"
date: "20/05/2021"
output:
  pdf_document: default
  html_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Repurposing Trial Data to Predict Death with AKI
We attempt to model the risk of death at 14 days with AKI using physiological & clinical variables collected as part of a clinical trial 

```{r}
library(tidyverse)
library(tidymodels)
#library(rms)
library(readr)
library(themis)
AKI_data <- read_csv("ELAIA-1_deidentified_data_10-6-2020.csv")

# Select only a subset of columns that will be used for modelling  
# Ensure these are either 'physiological' or from basic blood tests
AKI_set <- AKI_data %>%
  select(age,
         baseline_creat,
         bicarbonate_at_rand,
         creat_at_rand,
         death14,
         diastolic_at_rand,
         hemoglobin_at_rand,
         plateletcount_at_rand,
         potassium_at_rand,
         pulse_at_rand,
         resp_at_rand,
         sex,
         sodium_at_rand,
         systolic_at_rand,
         wbcc_at_rand
         )

```

## Exploratory Data Analysis
To look at the shape of the relationship of predcitors to death at 14 days to capture any non-linearity  
Create function the Hmisc spike lines & examine them
```{r}

#####
##Examine the relationships between the variables and outcome  
#####

# need rlang for NSE in following function
library(rlang)

# write function for plots
test.plots <- function(mydf, x, y, formula) {
  ggplot(mydf, aes(x = {{x}}, y = {{y}})) +  
  Hmisc::histSpikeg(formula = formula,
                    lowess = TRUE, data = mydf)
}

# Put them on a single plot
theme_set(theme_light())
age <- test.plots(AKI_set, age, death14, death14~age)
bicarb <- test.plots(AKI_set, bicarbonate_at_rand, death14, death14~bicarbonate_at_rand)
creat <- test.plots(AKI_set, creat_at_rand, death14, death14~creat_at_rand)
dbp <- test.plots(AKI_set, diastolic_at_rand, death14, death14~diastolic_at_rand)
hb <- test.plots(AKI_set, hemoglobin_at_rand, death14, death14~hemoglobin_at_rand)
plt <- test.plots(AKI_set, plateletcount_at_rand, death14, death14~plateletcount_at_rand)
kplus <- test.plots(AKI_set, potassium_at_rand, death14, death14~potassium_at_rand)
pulse <- test.plots(AKI_set, pulse_at_rand, death14, death14~pulse_at_rand)
resp <- test.plots(AKI_set, resp_at_rand, death14, death14~resp_at_rand)
Naplus <- test.plots(AKI_set, sodium_at_rand, death14, death14~sodium_at_rand)
sbp <- test.plots(AKI_set, systolic_at_rand, death14, death14~systolic_at_rand)
wbc <- test.plots(AKI_set, wbcc_at_rand, death14, death14~wbcc_at_rand)

initial_look <-
  cowplot::plot_grid(age, bicarb, creat, dbp, hb, plt, kplus, pulse, resp, Naplus, sbp, wbc,  nrow = 4, ncol = 3, scale = .9, labels = "AUTO")
```
## Modelling
Proceed via basic schema of:  
*?calculate sample size req'd
*split train/test 
*deal with imbalance
*Preprocess
*Train
*Metrics
```{r}
## Ensure death14 is a factor for outcome
AKI_mod <- AKI_set %>%
  mutate(death14 = factor(death14, 
                          levels = c(0, 1),
                          labels = c("alive", "dead")))

## Split into training & testing data
set.seed(19)
splits <- initial_split(AKI_mod, 
                        prop = 0.75, 
                        strata = death14)
AKI_train <- training(splits) 
AKI_test <- testing(splits)
 
# Create a recipe for the model 
lr_rec <- recipe(death14 ~ ., data = AKI_train) %>%
  themis::step_downsample(death14) %>% # to address class imbalance
  step_corr(all_numeric(), threshold = 0.7) %>% # remove vars that are highly correlated with others
  step_center(all_numeric()) %>%
  step_nzv(all_predictors()) # remove near zero variance predictors 

# Choose a model and an 'engine'
lr_mod <- logistic_reg(mode = "classification") %>%
  set_engine("glm")

# Combine into a workflow
lr_wflow <- workflow() %>%
  add_recipe(lr_rec) %>%
  add_model(lr_mod)

# train the model
AKI_fit_lr <- lr_wflow %>%
  fit(AKI_train)

# Predict on the test data
AKI_pred_lr <- predict(AKI_fit_lr, AKI_test) %>% 
  bind_cols(AKI_test %>% select(death14)) %>% 
  bind_cols(predict(AKI_fit_lr, AKI_test, type = "prob"))
# Check metrics
AKI_pred_lr %>% 
  metrics(truth = death14, .pred_class)

#ROC curve & AUROC 
AKI_pred_lr %>%
  roc_curve(truth = death14, .pred_alive)%>%
  autoplot()

AKI_pred_lr %>%
  roc_auc(truth = death14, .pred_alive)

#Calibration Plot
calibration_plot <- AKI_pred_lr %>% 
  mutate(
    alive = if_else(death14 == "alive", 1, 0),
    pred_rnd = round(.pred_alive, 2)
    ) %>% 
  group_by(pred_rnd) %>% 
  summarize(
    mean_pred = mean(.pred_alive),
    mean_obs = mean(alive),
    n = n()
    ) %>% 
  ggplot(aes(x = mean_pred, y = mean_obs)) +
  geom_abline(linetype = "dashed") +
  geom_point(aes(size = n), alpha = 0.5) +
  theme_light()
```


## Scrap 
```{r}
library(rms)
m0 <- lrm(death14 ~ 
            rcs(age, 4) +
            rcs(bicarbonate_at_rand, 4) +
            rcs(creat_at_rand, 4) +
            rcs(diastolic_at_rand, 4) +
            rcs(hemoglobin_at_rand, 4) +
            rcs(plateletcount_at_rand, 4) +
            rcs(potassium_at_rand, 4) +
            rcs(pulse_at_rand, 4) +
            rcs(resp_at_rand, 4) +
            rcs(sodium_at_rand, 4) +
            rcs(systolic_at_rand, 4) +
            rcs(wbcc_at_rand, 4)
          ,
          data = AKI_set, 
          x = TRUE, 
          y = TRUE
          )
```



